# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
library(data.table)
activity <- fread(unzip("activity.zip"))
```

## What is mean total number of steps taken per day?

```{r}
steps.per.day <- activity[, sum(steps, na.rm = T), by=date]$V1
hist(steps.per.day, breaks=15)
```

the mean **`r sprintf("%.1f",mean(steps.per.day))`** and median **`r median(steps.per.day)`** total number of steps taken per day

```{r}
steps.per.day <- aggregate(steps ~ date, activity, sum)$steps
hist(steps.per.day, breaks=15)
```

the mean **`r sprintf("%.1f",mean(steps.per.day))`** and median **`r median(steps.per.day)`** total number of steps taken per day

## What is the average daily activity pattern?

```{r}
d <- activity[, mean(steps, na.rm = T), by=interval]
setnames(d, "V1", "daily.activity")
with(d, plot(interval, daily.activity, type = "l"))

setkey(d, interval)
```

5-minute interval **`r d[daily.activity == max(d$daily.activity), interval]`**, on average across all the days in the dataset, contains the maximum number of steps


## Imputing missing values

**`r sum(is.na(activity$steps))`** is the total number of missing values in the dataset 


```{r}
DT <- copy(activity)
setkey(DT, interval)
DT <- merge(DT, d)

DT <- DT[, bb:= DT[, is.na(steps)] * d[, daily.activity]]

#DT <- DT[, bb:= DT[, is.na(steps)] * mean(steps.per.day)]
DT[is.na(steps), "steps"] <- DT[is.na(steps), as.integer(round(bb, 0))]

steps.per.day <- DT[, sum(steps), by=date]$V1

hist(steps.per.day, breaks=15)
```

the mean **`r sprintf("%.1f",mean(steps.per.day))`** and median **`r median(steps.per.day)`** total number of steps taken per day

## Are there differences in activity patterns between weekdays and weekends?
```{r}
DT$date <- as.Date(DT$date, format="%Y-%m-%d")
DT <- DT[, we := strftime(DT$date, "%u")>5]
DT$we <- factor(DT$we, levels=c(TRUE,FALSE), labels=c("weekend", "weekday"))

DT <- aggregate(steps ~ interval + we, DT, mean)

library(lattice)
with(DT, xyplot(steps~interval|we, 
     main="activity patterns between weekdays and weekends", 
     type = "l", 
     layout=c(1,2)))
```